---
title: "Class6 R Functions"
author: "Xiaoxue Lin"
date: "1/25/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## File reading (again)

Here we try to use **read.table()** and friends to input some example data into R

# Choose "Insert" from above and select R to input code type R to the script (Option + Command +I), shown below

Let's insert a code chunk.

```{r}
read.table("https://bioboot.github.io/bggn213_S18/class-material/test1.txt", header = TRUE, sep = ",")
```

```{r}
file1 <- "https://bioboot.github.io/bggn213_S18/class-material/test1.txt"
read.csv(file1)
```

```{r}

```

```{r}
file2 <- "https://bioboot.github.io/bggn213_S18/class-material/test2.txt"
read.csv(file2, sep = "$")
```

```{r}
file3 <- "https://bioboot.github.io/bggn213_S18/class-material/test3.txt"
data3 <- read.table(file3, header = FALSE, sep = "")
```

## R functions

```{r}
add <- function(x, y=1) {
# Sum the input x and y 
                x+y
}
```

# when playing this chunk, the function should appear under global environment

Lets use the **add()** function

```{r}
add(1)
```

```{r}
add(1,5)
```

```{r}
add(c(1,2,3,4))
```
```{r}
add(c(1,2,3,4), 4)
```

```{r}
## add(1, 2, 2) will have an error message because there are only two vectors but you are adding three
```


```{r}
## add(x = 1, y = "b") will have an error message because b is not a numerical
```

# Our 2nd function
```{r}
rescale <- function(x) {
  rng <-range(x)
  (x - rng[1]) / (rng[2] - rng[1])}
```

```{r}
rescale(1:10)
```

```{r}
# How would you get your function to work here...
rescale( c(1,2,NA,3,10) )
```
```{r}
rescale2 <- function(x){
  rng <-range(x, na.rm=TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}
```

```{r}
rescale2( c(1,2,NA,3,10))
```

```{r}
rescale2( c(1,2,NA,3,10))
```

```{r}
rescale3 <- function(x, na.rm=TRUE, plot=FALSE) {
   if(na.rm) {
     rng <-range(x, na.rm=na.rm)
   } else {
     rng <-range(x)
   }
   print("Hello")
   answer <- (x - rng[1]) / (rng[2] - rng[1])
   print("is it me you are looking for?")
   if(plot) {
      plot(answer, typ="b", lwd=4)
     print("please done ever sing again!")
}
   print("I can see it in ...")
}
```

```{r}
rescale3( 1:10 , plot=TRUE)
```

## Using the bio3d package

```{r}
library(bio3d)
```

```{r}
pdb <- read.pdb("1hbs")
```

What is in this **pdb** thing?

```{r}
pdb
```

```{r}
# can you improve this analysis code
library(bio3d)
s1 <- read.pdb("4AKE")  # kinase with drug
s2 <- read.pdb("1AKE")  # kinase no drug
s3 <- read.pdb("1E4Y")  # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

# plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
# plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
# plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
# 
# hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
# plot(hc)
```


```{r}


rbind(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
dist(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
hclust(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
```

