---
title: "Xiaoxue_Lin_Class6HW"
author: "Xiaoxue Lin"
date: "2/1/2019"
output:
  html_document: default
  pdf_document: default
  github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## to work with any dataset of protein structure..

The input of the function reads protein pdb files. 
The Output is the B factor plot of protein.
The following function first takes pdb of a protein with **read.pdb**, then uses **trim.pdb** to focus on chain A and "CA"" atom type. **plotb3** will plot the protein-drug interaction of selected sequence with B factor, and gives the output of one or more than one plots.

```{r}
library(bio3d)
# name "plot"" as the function
plot <- function (x)  {
  for (protein in x) {
  # s call s as read.pdb
  s <- read.pdb(protein)
  # trim.pdb produces a new smaller PDB object (protein chains) with a subset of atoms from a given larger PDB object.
  # call s.chainA to use trim.pdb function to select chain A 
  s.chainA <- trim.pdb(s, chain="A", elety="CA")
  # s.b places amino acid in **B factor**
  s.b <- s.chainA$atom$b
  # plot the data with amino acid on x axis and B factor (amino acid mobility) on y axis
  plotb3(s.b, sse=s.chainA, type = "l", ylab = "Bfactor")
  }
}
plot(c("4AKE", "1AKE", "1E4Y"))
```



