---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The PDB database

Let's examine the PDB database composition

```{r}
stats <- read.csv("Tubulin_acetyltransferase.csv", row.names = 1)
stats
```

Determine the percentage of structures solved by X-ray and Electron Microscopy

```{r}
pre.by.method <- stats$Total/sum(stats$Total) * 100

#Add names to each row of the stats

names(pre.by.method) <- rownames(stats)
names(pre.by.method)
```

What proportion of structures are protein

```{r}
portion.protein <- sum(stats$Proteins) / sum(stats$Total)
portion.protein
```

Paste with datapasta: use **Addins** to paste **as data.frame**

```{r}
test <- data.frame(stringsAsFactors=FALSE,
   Experimental.Method = c("X-Ray", "NMR", "Electron Microscopy", "Other",
                           "Multi Method", "Total"),
              Proteins = c(124770, 10988, 2057, 250, 127, 138192),
         Nucleic.Acids = c(1993, 1273, 31, 4, 5, 3306),
    ProteinNA.Complex = c(6451, 257, 723, 6, 2, 7439),
                 Other = c(10, 8, 0, 13, 1, 32),
                 Total = c(133224, 12526, 2811, 273, 135, 148969)
)

```

Read a pdb file

```{r}
library(bio3d)
```


```{r}
pdb <- read.pdb("1hsg")
pdb
```

From the above data info, it shows the number of chains (2 chains), non-protein molecules (127 residues of water, 1 residue of MK1 [the drug name])

Install a package from BitBucket: the 'devtools' package allows us to install development versions

```{r}
install.packages("devtools")
```

# Install the bio3d.view package from bitbucket

```{r}
devtools::install_bitbucket("Grantlab/bio3d-view")
```

To check the structure quickly instead of opening VMD to look at the structure

```{r}
library("bio3d.view")
# view(pdb, "overview", col="sse")
```

Extract the protein-only portion of this PDB structure and write it out to a new PDB file.

Extract the ligand (i.g. drug) and write it out to a new PDB file

```{r}
# Select all C-alpha atoms (return their indices)
ca.inds <- atom.select(pdb, "calpha")
ca.inds

# Select all drug molecules (return their indices)
drug.inds <- atom.select(pdb, "ligand")
drug.inds
```

In the above example, XYZ indices are three times larger than Atom Indices.

```{r}
# drug.inds$atom

# to access the atoms of the pdb file
pdb$atom

# to access individual columns
pdb$atom$chain

# to access first several columns
pdb$atom[1:3, 1:5] #first three atoms, and first five columns

# to access drug molecules
pdb$atom[drug.inds$atom, ]
```

To trim down the pdb to select specific molecules

```{r}
drug.trim <- trim.pdb(pdb, drug.inds) # drug.inds contains drug molecules that we obtained from atom.select() function
drug.trim
```

To write out the pdb file using write.pdb function

```{r}
write.pdb(drug.trim, file = "ligand.pdb")
```

Another example exercise

```{r}
pdb <- read.pdb ("1HEL")
pdb

# Normal Mode Analysis (details in next class)

modes <- nma(pdb)
```

