---
title: "K-means"
author: "Xiaoxue Lin"
date: "2/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```

```{r}
km <- kmeans (x, centers = 2, nstart = 20)
km
```

```{r}
km$size
```

```{r}
km$cluster
```

```{r}
plot(x, col=km$cluster)
```

```{r}
plot(x, col=km$cluster)
points(km$centers, col = "blue", pch=15, cex=2)
```

```{r}
# First we need to calculate point (dis)similarity
#   as the Euclidean distance between observations
dist_matrix <- dist(x)
# The hclust() function returns a hierarchical
#  clustering model
hc <- hclust(d = dist_matrix)
# the print method is not so useful here
hc
```

```{r}
plot(hc)
abline(h=6, col="red")
grp2 <- cutree(hc, h=6)
```

```{r}
plot(x, col=grp2)
```

```{r}
plot(hc)
abline(h=2.5, col="blue")
grp6 <- cutree(hc, h=2.5)
# alternatively, to cut tree, use cutree(hc, k=2)
table(grp6)
```

```{r}
plot(hc)
grp3 <- cutree(hc, k = 3)
plot(x, col=grp3)
```

# Using different hierarchical clustering methods

```{r}
plot(hc)
hc.complete <- hclust(dist(x), method="complete")
plot(hc.complete)
hc.single <- hclust(dist(x), method = "single")
plot(hc.single)
hc.centroid <- hclust(dist(x), method = "centroid")
plot(hc.centroid)
```

```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
  matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2),   # c1
  matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
  matrix(c(rnorm(50, mean = 1, sd = 0.3),           # c3
           rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
# Step 2. Plot the data without clustering
plot(x)
# Step 3. Generate colors for known clusters
#         (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```

```{r}
distx <- dist(x)
hcluster <- hclust(distx)
plot(hcluster)
```

```{r}
cut_hcluster <- cutree(hcluster, k = 3)
plot(x, col = cut_hcluster)
```

# PCA: principal Component Analysis

```{r}
## You can also download this file from the class website!
mydata <- read.csv("https://tinyurl.com/expression-CSV",
row.names=1)
head(mydata)
```

## lets do PCA with prompt() function
**NOTE** prcomp() expects the samples to be rows and genes to be columns so we need to first transpose the matrix with the t() function!

```{r}
#PCA in R expect the samples to be on top of column, so use t()
head(t(mydata)) 
pca <- prcomp(t(mydata), scale=TRUE)
summary(pca)
```
 
 make our first PCA plot
 
```{r}
dim(pca$x)
plot(pca$x[,1],pca$x[,2], xlab = "PC1", ylab = "PC2")
#look for the split mostly on PC1 axis to see how separated these two groups are
```
 
```{r}
## Precent variance is often more informative to look at
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 3)
pca.var.per
```
 
 make a scree plot
 
```{r}
barplot(pca.var.per, main="Scree Plot",
         xlab="Principal Component", ylab="Percent Variation")
```
 
 lets make the plot a bit nicer
 
```{r}
 ## A vector of colors for wt and ko samples
## reads from the middle of the parenthesis, run in console
 colvec <- as.factor( substr( colnames(mydata), 1, 2) )
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
xlab=paste0("PC1 (", pca.var.per[1], "%)"), 
      #paste0 makes a string from the data
ylab=paste0("PC2 (", pca.var.per[2], "%)"))
```
 
 
 Make UK food plot data
 
```{r}
#read data
uk<-read.csv("UK_foods.csv")
uk
```

To access the name of the first row
 
```{r}
uk[,1]
#use uk[,-1] to access every row except the first row
```
 
 To read the first row properly in the first place
```{r}

#make the first column as row name, use **row.names = 1**

x <- read.csv("UK_foods.csv", row.names = 1) 
x
```
 
 Dimensions of the table
 
```{r}
dim(x)
dimension <- c(nrow(x), ncol(x))
dimension
```
 
Preview the first 6 rows and last 6 rows
```{r}
head(x)
tail(x)
```

to draw a pair plot

```{r}
pairs(x, col=rainbow(10), pch=16)
```

First row is England data plotted against other countries in x axis. Same for the next rows

Look at the summary of PCA

```{r}
pca <- prcomp(t(x))
pca
summary(pca)
```

to make a PCA plot

```{r}
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x))
```

to color the countries

```{r}
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500), col="gray")
text(pca$x[,1], pca$x[,2], colnames(x) ,col=c("orange","red","blue","darkgreen"))
```



```{r}
## Lets focus on PC1 as it accounts for > 90% of variance 
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )
```
