---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the Bio3D package and use the get.pdb() function to download the "1hsg" PDB entry into your RStudio Project directory.

```{r}
library(bio3d)

pdb.code <- "1hsg"
file.name <- get.pdb(pdb.code)
```

use the read.pdb()function to read this PDB file into R so we can prepare it for further analysis. Write a new PDB format file.
We will also do the same for the bound ligand.

```{r}
hiv <- read.pdb(file.name)
hiv
```

From the above read.pdb function, there are 127 water molecules, 1 MK1 (drug compound)

Protein extraction. You could look at the function of **trim.pdb** to know how to extract protein.

```{r}
prot <- trim.pdb(hiv, "protein")
prot
```

Convert the extracted protein into a new file.
Use **paste()** to *call* the extracted protein with a new name, no files are created

```{r}
#Without [sep=], the file name would contain a space, which is NOT desirable
prot.filename <- paste(pdb.code, "_protein.pdb", sep="")
write.pdb(prot, file=prot.filename)
```

Doing the same thing for ligand

```{r}
mk1 <- trim.pdb(hiv,"ligand")
mk1
```

```{r}
lig.filename <- paste(pdb.code, "_ligand.pdb", sep="")
write.pdb(mk1, file=lig.filename)
```

To inspect the docking results, we will convert the **all.pdbqt** to a PDB format file

```{r}
library(bio3d)
res <- read.pdb("all.pdbqt", multi=TRUE)
write.pdb(res, "results.pdb")
```

Load the original 1hsg.pdb file and the new **results.pdb** file into VMD.

Use the **rmsd()** function to compare to your docking results with original ligand.

```{r}
# res <- read.pdb("all.pdbqt", multi=TRUE)
ori <- read.pdb("ligand.pdbqt")
rmsd(ori, res)
```

Lower value of RMSD (root mean square distance) means better drug compound.

Exploring the conformational dynamics of protein **1hel**

```{r}
library(bio3d)
pdb <- read.pdb("1HEL")
```

the function **nma()** is used perform the normal mode calculation to probe large scale mobility of biomolecules.

```{r}
modes <- nma(pdb)
plot(modes, sse=pdb)
```

Visualize NMA results

```{r}
 # Visualize NMA results
mktrj(modes, mode=7, file="nma_7.pdb")
```
