---
title: "Class 7 functions and packages"
author: "Xiaoxue Lin"
date: "1/30/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Functions revisited

load (i.e. **source**) our rescale() function from last day

```{r}
source("http://tinyurl.com/rescale-R")
```

test this function
```{r}
rescale(1:5)
```

```{r}
#' rescale(c(1:5, "string"))
```

We want to make this function more robust to these "string"" types of errors
two functions to use, warning() and stop()

```{r}
#rescale2(c(1:5, "string"))
```

```{r}
is.numeric(1:5)
```

```{r}
is.numeric("string")
```

```{r}
is.numeric(c(1:5, "string"))
```

```{r}
!is.numeric(1:5)
```

```{r}
x <- c( 1, 2, NA, 3, NA)
y<-c(NA,3,NA,3, 4)
```

```{r}
is.na(x)
is.na(y)
```

```{r}
sum(is.na (x) & is.na (y)) #this is a snippet
# which(is.na(x) & is.na(y)) indicates where in the vector is the true element
```

Now take our working snippet and make first function

```{r}
both_na <- function (x,y) {
  # Check for NA elements in both input vectors
  sum(is.na (x) & is.na (y))
}
```

```{r}
x <- c( NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)
y3 <- c( 1, NA, NA, NA, NA)
```

```{r}
# What will this return?
both_na(x,y2)
```

When comparing x with y2, the last NA with y2 has no paring, so the first element of x is added on to the end of x to pair with y2.

```{r}
both_na <- function (x, y) {
  if(length(x) != length(y)) {
stop("Input x and y should be vectors of the same length")
# be easily missed especially in scripts.
}
  na.in.both <- ( is.na(x) & is.na(y) )
  na.number  <- sum(na.in.both)
  na.which   <- which(na.in.both)
  message("Found ", na.number, " NA's at position(s):",
          paste(na.which, collapse=", ") )
  return( list(number=na.number, which=na.which) )
}
```

```{r}
# library("sigma")
# data <- system.file("examples/ediaspora.gexf.xml", package = "sigma")
# sigma(data)
```

```{r}
install.packages("ggplot2")
```

```{r}
install.packages("plotly")
```

```{r}
library(ggplot2)
library(plotly)
p <- ggplot(data = diamonds, aes(x = cut, fill = clarity)) +
            geom_bar(position = "dodge")
ggplotly(p)
```

