---
title: "Class9Project_BreastCancerData"
author: "Xiaoxue Lin"
date: "2/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


To download the file, use **read.csv()** function, and the file name should be parenthesized

```{r}
read.csv("WisconsinCancer.csv")
```


```{r}
#save the file to a new data directory
fna.data <- "WisconsinCancer.csv"

# Complete the following code to input the data and store as wisc.df
wisc.df <- read.csv(fna.data)
```

Double check data with **view()** or **head()**

```{r}
head(wisc.df)
```

The last column (x) is weird. Let's check how many patients and features are in this data. 

```{r}
nrow(wisc.df)
ncol(wisc.df)
```

Use **as.matrix** to convert the other features (i.e. columns) of the data (in columns 3 through 32) to a matrix. 

```{r}
wisc.data <- wisc.df[ ,3:32]
head(wisc.data)
```

Add patient ID as row names

```{r}
rownames(wisc.data) <- wisc.df$id
head(wisc.data)
```

If we only want to include two individual columns instead of the whole set of columns

```{r}
wisc.df[ , c(4, 10, 32)]
```

Q: How many malignant /  belign samples are there?

```{r}
table(wisc.df$diagnosis)
```

Q: How many of the columns have mean values?

```{r}
colnames(wisc.data)
grep("mean", x = colnames(wisc.data))

#To know how many "means"" there are
length(grep("mean", x = colnames(wisc.data)))
```

```{r}
inds <- grep("mean", x = colnames(wisc.data))
colnames(wisc.data)[inds]
```

apply() function will apply any function at a given position included in the argument

```{r}
apply(wisc.data, 2, mean) # 2 as column

# to round up the figures
round(apply(wisc.data, 2, mean), 2) 
```

Standard deviation
```{r}
round(apply(wisc.data, 2, sd), 2)
```

To compute a PCA

```{r}
wisc.pr <- prcomp(wisc.data, scale = TRUE)
summary(wisc.pr)
```

From the above data, PC1 proportion of variance is 44% of the entire number of samples. PC2 does not do as well as PC1 (second principal component). 






To do principal component analysis, first step is the see if the scale needs means and standard deviation

```{r}
colMeans(wisc.data)
```

Creating a plot using **biplot** for standard, simple data plot

```{r}
biplot(wisc.pr)
```

The data is really messy, we need to take PC1 and PC2 for plotting the data. One of the 5 elements in **wisc.pr** list is **x**, which is the PC data from PC1 to 30

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[, 2])
```

Let's color the plot using the expert diagnosis

```{r}
#wisc.df$diagnosis
plot(wisc.pr$x[,1], wisc.pr$x[, 2], col = wisc.df$diagnosis, 
     xlab = "PC1", ylab = "PC2")
```

Variance:
Calculate the variance of each principal component by squaring the sdev component of wisc.pr (i.e.  wisc.pr$sdev^2). Save the result as an object called pr.var.

```{r}
pr.var <- wisc.pr$sdev^2
head(pr.var)
```

Proportion of each variance in the total variance

```{r}
pve <- pr.var / sum(pr.var)
head(pve)
```

Round up the numbers
```{r}
round(pve * 100, 2)
```

create a plot of variance explained for each principal component
type="l" would give a line plot, "b" a dot+line plot

```{r}
plot(pve, ylim = c(0, 1), type="b")
```

Let's do a simple barplot

```{r}
# Alternative scree plot of the same data, note data driven y-axis
barplot(pve, ylab = "Precent of Variance Explained",
        axes = FALSE, # deplete axes labels
        names.arg = paste("PC", 1:length(pve)), las = 2) 
axis(2, at=pve, labels=round(pve, 2)) #2 is the y plot label, this will give back the yaxis label
```

Clustering in Principal Component Analysis
For hcluster we need a distance matrix and we get this from our PCA result (i.g. wisc.pr$x)

```{r}
d <- dist(wisc.pr$x[,1:2])
hc <- hclust(d, method = "ward.D2")
plot(hc)
```

To access the members under each branch of the tree plot, use **cutree()**

```{r}
cut <- cutree(hc, k = 3)
head(cut)
```

To know how many data in each group
```{r}
grp3 <- table(cutree(hc, k = 3))
head(grp3)
```



To plot the data via different cuts
```{r}
plot(wisc.pr$x[, 1:2],xlab = "PC1", ylab = "PC2", 
     col =cut)
```

Cross tabulation

```{r}
#table(grp3)
#First assign digits as diagnosis result: 0 for B, 1 for M
diagnosis <- wisc.df$diagnosis == "M"
table(cut, diagnosis)
```

```{r}
table(cut, wisc.df$diagnosis)
```

Predict new data from previous model

```{r}
#url <- "new_samples.csv"
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```

Plot the new data

```{r}
plot(wisc.pr$x[,1:2], col = wisc.df$diagnosis) 
points(npc[,1], npc[,2], col="blue", pch=16, cex = 3)
```

